---
title: |
  | \vspace{4cm} \LARGE{Data Mining Assignment}
subtitle: |
 Assignment 1
author: "Hadeer Elmahdy"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
header-includes: 
  - \renewcommand{\and}{\\}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage 
\tableofcontents 
\newpage 


## I. The Problem 

SafeBabies is a large company who is producing car seats for babies and toddlers. They sell their products all over the US and abroad. The management team has hired you as a Business Analytics consultant to help them maximize their profit. 

You have been told that the cost of producing each car seat is $55.0 

### Your primary task is to determine:
1-	The optimal price for selling the car seats at those stores where the shelve location is “good” (i.e., the product is highly visible)
2-	The optimal price for selling the car seats at those stores where the shelve location is “bad” (i.e., the product is poorly visible)

3-	Plot the optimal price for selling the car seats at those stores where the shelve location is “good” and separately for those stores where the shelve location is “bad” when varying the production costs from $40 to $85 (in $5 increments).




```{r}
#Loading Liberaries 
library(dplyr)
library(ISLR)
library(ggplot2)

```


## II. Data Prepration & Exploration

```{r , figures-side, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE}

SafeBabies <- Carseats %>% select("Sales", "Price", "ShelveLoc")
summary(SafeBabies)
head(SafeBabies)

#Check for missing values
table(is.na(SafeBabies))
sum(is.na(SafeBabies))  #no missing values

#Dividing dataset based on Shelve location
Good_Shelve <- filter(SafeBabies, ShelveLoc == "Good")

Bad_Shelve <- filter(SafeBabies, ShelveLoc == "Bad")


#EDA Plots
ggplot(SafeBabies) +
 aes(x = Sales) +
 geom_histogram(bins = 30L, fill = "#0c4c8a") +
 theme_minimal()

ggplot(SafeBabies) +
 aes(x = Price) +
 geom_histogram(bins = 30L, fill = "#0c4c8a") +
 theme_minimal()


ggplot(SafeBabies) +
 aes(x = ShelveLoc, y = Price) +
 geom_boxplot(fill = "#0c4c8a") +
 labs(x = "Shelve Location", title = "Price based on Shelve Location") +
 theme_minimal()


ggplot(SafeBabies) +
 aes(x = ShelveLoc, y = Sales) +
 geom_boxplot(fill = "#0c4c8a") +
 labs(x = "Shelve Location", title = "Sales based on Shelve Location") +
 theme_minimal()


ggplot(SafeBabies) +
 aes(x = Price, y = Sales) +
 geom_point(size = 1L, colour = "#0c4c8a") +
 labs(title = "Sales based on Price") +
 theme_minimal()

```


#Summary: Price is slightly higher for car seats with good shelve location.
#Intuitively, Sales are higher for car seats with good shelve location.
#There is a negative relationship between Sales and Price.


\newpage

## III. Building the Model
  Building linear model to predict sales for car seats with good and bad shelve location separately

```{r}
Model_GoodShelve <- lm(Sales ~ Price, data = Good_Shelve)

summary(Model_GoodShelve)

Model_BadShelve <- lm(Sales ~ Price, data = Bad_Shelve)

summary(Model_BadShelve)


#Both models confirm the negative relationship between Sales and Price. As price increases, sales decrease. 
#R squared for the good-shelve model is 43%, which means that the model explains 43% of the variability in the price of car seats with good shelve location. R squared for the bad-shelve model is 30, meaning that model explains only 30% of the variability in the price of car seats with bad shelve location. Sales= b1*price + b0 

```

## IV. Problem Formulation:

Total Profit= Sales * Unit Profit
Unit Profit= Unit price – Unit cost     
Sales= b1*price + b0
Total Profit= (Unit price – Unit cost) *(b1*price+b0) 
            = b1*price^2+(b0 -cost*b1)*price -cost*b0
            
Now we will set the derivative to Zero to find optimal price
           
Total Profit derivative = 2b1*price + b0-b1*Cost

Then we solve for the price:

Price = (-b0)-(b1*Cost)) / 2b1


## V.  Problem Solving:
```{r}
#create function to calculate optimal price for car seats given fixed productions cost of $55
Optimal_Price <- function(cost, b0, b1) {
    return(((-b0) + (b1 * cost)) / (2 * b1))
}


#Optimal price for car seats with good shelve location given that the cost = $55 is $164.0731 
OptimalPrice_Good <- Optimal_Price(55, Model_GoodShelve$coefficients[1], Model_GoodShelve$coefficients[2])
OptimalPrice_Good


#Optimal price for car seats with bad shelve location given that the cost = $55 is $ 134.6435 
OptimalPrice_bad<- Optimal_Price(55, Model_BadShelve$coefficients[1], Model_BadShelve$coefficients[2])
OptimalPrice_bad

```


### Plot the optimal price when cost varies between $40 and $85

```{r}
Production_Costs <- seq(40,85, by =5) #production costs from $40 to $85 (in $5 increments)


for (i in Production_Costs) {
  Optimal_good <- Optimal_Price(Production_Costs, Model_GoodShelve$coefficients[1], Model_GoodShelve$coefficients[2])
  Optimal_bad <- Optimal_Price(Production_Costs, Model_BadShelve$coefficients[1], Model_BadShelve$coefficients[2])
}


#combining production costs and optimal prices for good and bad shelve location in a data frame 
df <- as.data.frame (cbind(Production_Costs, Optimal_good, Optimal_bad))
df


#Plotting Optimal Prices for Production costs range $40:$85 -Bad Shelve location
ggplot(df) +
 aes(x = Production_Costs, y = Optimal_bad) +
 geom_line(size = 1L, colour = "#0c4c8a") +
 labs(y = "Optimal Price", title = "Optimal Price for Production costs range $40:$85 -Bad Shelve ") +
 theme_minimal()



#Plotting Optimal Prices for Production costs range $40:$85 -Good Shelve location
ggplot(df) +
 aes(x = Production_Costs, y = Optimal_good) +
 geom_line(size = 1L, colour = "#0c4c8a") +
 labs(y = "Optimal Price", title = "Optimal Price for Production costs range $40:$85 -Good Shelve") +
 theme_minimal()



```





